---
title: "Gov 50 Final Project"
author: "Kieran Kelly"
description: "International Law: An Analysis of Efficacy"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to international law. I am excited to leverage data to better understand why international legal institutions act in the ways they do. This may include the efficacy of the institutions, the actions taken by institutions, or other forms of analysis related to international law.

## Research Proposal

Are international legal agreements more enforceable when parties can negotiate and coordinate informally? In this study, I will analyze the relationship between the ability for international parties to negotiate in an informal manner and the enforceability of the agreement under which they would have such ability. I hypothesize that those agreements which allow for nations to coordinate in informal negotiations will be more effective in the metric of enforceability. My theory in support is rooted in the inherent power of the state; in agreements where nation states have more authority over the conditions with which they can interact with their peers, these conditions, with this greater leeway, will be more enforceable. My explanatory variable regards whether or not international legal agreements allow for informal dispute resolution. The outcome variable refers to whether there exists a disparity among agreements where this informal coordination is possible. In the data set, there is a binary variable which measures both of these ideas, so the analysis of both variables should be fairly straightforward. If international legal agreements where informal dispute resolutions are available tend to be more enforceable, this analysis would support my thesis that international legal agreements where informal coordination is allowed are more enforceable. This could take the form of a higher mean value in the enforceability variable among the group of agreements with informal dispute resolution options, skewed values on a dotplot, or other demonstrations of statistical correlation among these two variables. However, if this kind of relationship is not evident in the data, my hypothesis will be proven incorrect.

## First Data Visualization

```{r}
library(readr)
library(dplyr)
library(tidyverse)
```



```{r}

data_rep_ch7 <- read_csv("data_rep_ch7.csv")

informal_vs_enforce <- data_rep_ch7 |>
  group_by(info, enforce) |>
  summarise(n = n())


informal_vs_enforce <- informal_vs_enforce |>
  mutate(informal_power = if_else(info==1, "Informal Power", "No Informal Power"),
         enforce_power = if_else(enforce==1, "Not Enforceable", "Enforceable"))

informal_vs_enforce_barplot <- informal_vs_enforce |>
  ggplot(aes(fill= enforce_power,
             y = n,
             x=informal_power)) + geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Informal Conflict Resolution vs. Enforceability of International Agreements",
       x = "Informal Power Capabilities", y = "Amount of Agreements")
informal_vs_enforce_barplot
```

