---
title: "International Law: An Analysis of Efficacy"
author: "Kieran Kelly"
description: "Does the availability of informal dispute resolution capabilities within international agreements improve the enforceability of such agreements?"
output:
  distill::distill_article:
    self_contained: false
---

```{r include=FALSE}
#Options
options(
  scipen = 9,
  width = 100)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center")

library(readr)
library(dplyr)
library(tidyverse)

data_rep_ch7 <- read_csv("data_rep_ch7.csv")
```


## [Introduction]{.underline}

Are international legal agreements more enforceable when parties can negotiate and coordinate informally? In this study, I will analyze the relationship between the ability for international parties to negotiate i![]()n an informal manner and the enforceability of the agreement under which they would have such ability. I hypothesize that those agreements which al![]()low for nations to coordinate in informal negotiations will be more effective in the metric of enforceability. My theory in support is rooted in the inherent power of the state; in agreements where nation states have more authority over the conditions with which they can interact with their peers, these conditions, with this greater leeway, will be more enforceable.

## [Data]{.underline}

The data for this project is sourced from Barabara Koremenos's book *The Continent of International Law: Explaining Agreement Design*. In this book, Koremenos explains the ins and outs of international agreements and argues that the nuts and bolts matter. The inputs to these agreements, including the design of the agreement, the matter for which the agreement deals, and the amount of parties signing onto the agreement, impact the efficacy of such agreement. 

Each chapter of Koremenos's book includes a different data set. My data set comes from her seventh chapter, "Dispute resolution provisions," in which she analyzes what aspects of international agreements contribute to better resuslts in dispute resolution. The data studied is derived from a cross-sectional data set of a large swath of international agreements and varying characteristics of such agreements. In my analysis, I specifically analyze two binary variables: "enforce" and "info." In her code, "enforce" is coded as a '1' if there is a problem with enforcement and '0' if informal power may be used in dispute resolution. This informal power may refer to side payments, fines, or other kinds of compensation that may resolve disputes without formal and traditional negotiations.

My explanatory variable regards whether or not international legal agreements allow for informal dispute resolution. The outcome variable refers to whether there exists a disparity among agreements where this informal power allotted to the agreement exists. Below is a barplot that shows the quantity of agreements that are enforceable or not enforceable within the data set.

```{r}


informal_vs_enforce <- data_rep_ch7 |>
  group_by(info, enforce) |>
  summarise(n = n())


informal_vs_enforce <- informal_vs_enforce |>
  mutate(informal_power = if_else(info==1, "Informal Power", "No Informal Power"),
         enforce_power = if_else(enforce==1, "Not Enforceable", "Enforceable"))

enforceability <- data_rep_ch7 |>
  group_by(enforce) |>
  summarise(n = n())

enforceability <- enforceability |>
  mutate(enforce_new = if_else(enforce==1, "Not Enforceable", "Enforceable"))

enforce_barplot <- enforceability |>
  ggplot(aes(x = enforce_new,
             y = n)) + 
  geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Enforceable vs. Not Enforceable International Agreements",
       x = "Enforceability of Agreements",
       y = "Amount of Agreements")
enforce_barplot
```

## [Results]{.underline}

```{r}
informal_vs_enforce_barplot <- informal_vs_enforce |>
  ggplot(aes(fill= enforce_power,
             y = n,
             x=informal_power)) + geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Informal Conflict Resolution vs. Enforceability of International Agreements",
       x = "Informal Power Capabilities", y = "Amount of Agreements")
informal_vs_enforce_barplot

```

```{r}
enforceable_informal_regression <- lm(enforce ~ info + arb + med, data = data_rep_ch7)
summary(enforceable_informal_regression)$coefficients |>
  knitr::kable(digits = 2)
```

The barplot above shows the quantity of international agreements that fall within four different categories: those that allow for informal dispute resolution, and whether or not they are enforceable, and those that do not allow for informal dispute resolution, and whether they are enforceable. We see that there is a convincingly larger amount of enforceable agreements that do not allow for informal power than enforceable agreements that do allow for informal power. The quantity of agreements that are not enforceable are relatively similar for both those that contain informal dispute resolution power and those that lack such power, though there are slightly more agreements that are not enforceable that allow for informal dispute resolution.

The relationship is formalized in the multivariate regression. The regression shows us that with the allowance of informal dispute resolution power, 8% more agreements show that they are not enforceable than those that do not allow for informal dispute resolution. The regression also shows even stronger relationships between the ability for disputes to be resolved through arbitration and mediation, with 34% and 19% more agreements having an enforcement problem with these kinds of dispute resolution methods, respectively.

We can not conclude that the 'info' variable has a causal effect on the 'enforce' variable based on this regression. This is because the p-value reported is .30 which would imply that with significance levels of .01, .05, and .1, the relationship would still be statistically insignificant.

## Conclusion

This result is contrary to what I predicted in my hypothesis. The data indicates that there is a marginally negative relationship between the availability of informal dispute resolution tactics, such as side payments or other forms of informal negotiations, and the enforceability of such agreements. From this, we can conclude that agreements are slightly correlated with higher enforceability in cases when the ways in which nations can resolve disputes is stricter. The analysis contributing to my hypothesis was based on the premise that with informal dispute resolutions, nations are given more leeway in their ability to resolve disputes. However, through the analysis of the data, we can assume that with stricter guidelines, agreements are, too, marginally more enforceable. It is also important to remember that different agreements have different characteristics. As we can see in the regression, arbitration capabilities, mediation capabilities, as well as the issue with which the agreement deals are all impactful regarding the enforceability of the agreement. The regression implies an even stronger relationship between arbitration and mediation capabilities and the enforcement of the agreement, relative to the availability of informal resolution tactics. By analyzing a larger swath of international agreements, and more importantly, groups of agreements with similar varying characteristics, we can more accurately study the relationship between informal dispute resolution tactics and the enforceability of such agreements. By keeping these other variables constant, we can more accurately study this relationship by minimizing potential threats to inference. 
